\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{accents}


% Copyright
%\setcopyright{none}
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


% DOI
\acmDOI{xx.xxx/xxx_x}

% ISBN
\acmISBN{978-1-4503-8104-8/21/03}

%Conference
\acmConference[SAC'21]{ACM SAC Conference}{March 22-March 26, 2021}{Gwangju, South Korea}
\acmYear{2021}
\copyrightyear{2021}


\acmArticle{4}
\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}


\begin{document}
\title{Serendipity Oriented and Ontology-based Recommender System for E-tourism}
%\title{RECESO: Recommender System for E-tourism with Serendipity and Ontology-based}
%\titlenote{Produces the permission block, and
 % copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
 % \texttt{acmart.pdf} document}
 

%\author{Gustavo Castellanos}
%\orcid{1234-5678-9012}
%\affiliation{%
%  \institution{Universidad Simón Bolívar}
%  \city{Caracas} 
%  \state{Venezuela} 
 % \postcode{1080-A}
%}
%\email{14-10192@usb.ve}

%\author{Yudith Cardinale}
%\affiliation{%
%  \institution{Universidad Simón Bolívar, Caracas, Venezuela}
%  \institution{Universidad Católica San Pablo, Arequipa, Perú}
%  \city{Caracas} 
%  \state{Venezuela} 
 % \postcode{1080-A}
%}
%\email{ycardinale@usb.ve}

%\author{Philippe Roose}
%\affiliation{%
 % \institution{Université de Pau et des Pays de l’Adour}}
 % \streetaddress{LIUPPA -- T2I64600}
 % \city{Anglet} 
 % \country{France}}
% \email{philippe.roose@iutbayonne.univ-pau.fr}


% The default list of authors is too long for headers}
%\renewcommand{\shortauthors}{G. Castellanos, Y. Cardinale, P. Roose.}


\begin{abstract}
Frequently, travelers try to 
%manually 
collect 
%enough 
information for planing a trip
%to go to a destination 
or when being at 
%this 
the destination.
%, but 
This task may be overwhelming due to the big amount of available information sources
%options 
(including local apps download). Usually, tourists depend on places' reviews to make the choice, but this implies prior knowledge of the 
%existence of the 
touristic places. These reviews are often requested on tourism sites and applications, most of which follow the PULL paradigm, i.e., the user has to explicitly search for suggestions
%/recommendations 
through interaction with the application. 
%'s graphical user interface. 
In contrast,
%On the other hand, 
a
PUSH approach, in which the application proactively triggers a recommendation process when
necessary -- e.g.,
%%for example, 
%when 
the user is close to a point of interest  
%(POI) 
or  
when lunch time is approaching --
%, or
%when weather conditions change -- 
seems to be a more reasonable solution. 
%Having said that, 
In this context, 
%in this work 
we propose RECESO, a user-centric 
%hybrid 
recommender system that supports both PULL and PUSH approaches. We propose
%using 
an ontology-based spreading activation algorithm for %content-based and 
context-aware recommendations, with a focus on 
decreasing predictable outputs 
%under same situations, i.e., 
and increasing serendipity  based on an {\it aging-like} approach. To demonstrate its suitability and performance, we test it in different simulated scenarios, varying users' profiles and preferences and  context parameters.
%using different 
%weights or priorities for the predicted preferences, context activation, aging, and distances of the recommended places. 
Results show how the spreading activation and the proposed aging system provide
%give 
relevant and varied recommendations, according to users' preferences,
%for the users, 
while taking into account their context. We show
%, and 
how the system 
%could be 
is configured to focus more on either relevant, varied, context-activated, or near places.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%


\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002951.10003227.10003233.10003449</concept_id>
<concept_desc>Information systems~Reputation systems</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003120</concept_id>
<concept_desc>Human-centered computing</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Reputation systems}
\ccsdesc[500]{Human-centered computing}


\keywords{Tourism, Recommender Systems, Context Awareness, Ontology, User-centric}


\maketitle

\input{Introduction}
\input{Related}
\input{Proposal}
\input{StudyCase}
\input{Conclu}

\vspace{-0.2cm}
\bibliographystyle{ACM-Reference-Format}
\balance
\bibliography{bibliography} 

\end{document}
